---
title: "STA7734 Project"
author: "Yuan Du"
date: "10/2/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(dplyr)
library(ggplot2)
library(psych)
library(corrplot)
```

## read data
```{r read}
Data <- read_csv("C:/Work/Project/STA7734/SVM-Asymptotic-Normality/Data/data.csv",
                 col_names = T,
                 col_types=cols(ID =col_character()，
                                Diagnosis=col_character(),
                                "3" = col_double(),
                                "4" = col_double(),
                                "5" = col_double(),
                                "6" = col_double(),
                                "7" = col_double(),
                                "8" = col_double(),
                                "9" = col_double(),
                                "10" = col_double(),
                                "11" = col_double(),
                                "12" = col_double(),
                                "13" = col_double(),
                                "14" = col_double(),
                                "15" = col_double(),
                                "16" = col_double(),
                                "17" = col_double(),
                                "18" = col_double(),
                                "19" = col_double(),
                                "20" = col_double(),
                                "21" = col_double(),
                                "22" = col_double(),
                                "23" = col_double(),
                                "24" = col_double(),
                                "25" = col_double(),
                                "26" = col_double(),
                                "27" = col_double(),
                                "28" = col_double(),
                                "29" = col_double(),
                                "30" = col_double(),
                                "31" = col_double(),
                                "32" = col_double()
                                )) %>%
                        rename( "radius" = "3",
                                "texture" = "4",
                                "perimeter" = "5",
                                "area" = "6",
                                "smoothness" = "7",
                                "compactness" = "8",
                                "concavity" = "9",
                                "concave_points" = "10",
                                "symmetry" = "11",
                                "fractal_dimension" = "12",
                                "radiusSE" = "13",
                                "textureSE" = "14",
                                "perimeterSE" = "15",
                                "areaSE" = "16",
                                "smoothnessSE" = "17",
                                "compactnessSE" = "18",
                                "concavitySE" = "19",
                                "concave_pointsSE" = "20",
                                "symmetrySE" = "21",
                                "fractal_dimensionSE" = "22",
                                "radiusW" = "23",
                                "textureW" = "24",
                                "perimeterW" = "25",
                                "areaW" = "26",
                                "smoothnessW" = "27",
                                "compactnessW" = "28",
                                "concavityW" = "29",
                                "concave_pointsW" = "30",
                                "symmetryW" = "31",
                                "fractal_dimensionW" = "32")

spec(Data)
```


## Check Data and missing values

```{r check, echo=FALSE}
head(Data)
tail(Data)
sapply(Data, function(x) sum(is.na(x)))
```
## Data visiualization
```{r}

psych::describe(Data)
Data_drop <- Data[,-1] #drop ID
psych::describeBy(Data_drop,Data_drop$Diagnosis)

#Plot Diagnosis 
ggplot(Data, aes(Diagnosis)) +
  geom_bar()

#Plot histogram for all Variables
Data_drop %>%
  keep(is.numeric) %>% 
  gather() %>% 
  ggplot(aes(value)) + 
    geom_histogram() + 
    facet_wrap(~key, scales = "free")
#Plot histogram for all Variables by Diagnosis
Data_drop %>%
  gather(key = "Var", value = "value", -Diagnosis) %>% 
  ggplot(aes(x=value, fill=Diagnosis)) + 
    geom_histogram(bins = 60) + 
    facet_wrap(~Var, scales = "free")
```

## Correlaion analysis by spearman since most variables are non normal distributed
```{r}
#create correlation matrix
Cor <- Data_drop %>% 
       select(-Diagnosis) %>%
       cor(method="spearman")

corrplot(Cor, type="upper", order="hclust")
```

## Data Transformation/ Dimension reduction by using PCA
```{r}
Data_pca <- prcomp(Data_drop[,-1], center=TRUE, scale=TRUE)

plot(Data_pca, type="l", main='')
grid(nx = 10, ny = 14)
title(main = "Principal components weight", sub = NULL, xlab = "Components")
box()
summary(Data_pca)
```
## visualize which variables are the most influential on the first 2 components

```{r}
pca_df <- as.data.frame(Data_pca$x)
ggplot(pca_df, aes(x=PC1, y=PC2, col=Data$Diagnosis)) + geom_point(alpha=0.5)
```

## Individuals with a similar profile are grouped together
```{r}
library(factoextra)
fviz_eig(Data_pca)
fviz_pca_var(Data_pca,
             col.var = "contrib", # Color by contributions to the PC
             gradient.cols = c("#00AFBB", "#E7B800", "#FC4E07"),
             repel = TRUE     # Avoid text overlapping
             )
```
Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
